---
title: "Introduction to R Programming"
author: "Muhammad Shahid Mirza"
date: "Jan 24, 2021"
time: "21:35"
output: html_document
---


##### NOTE: 
1. Change author name and date to your exercise submission date in above section
2. Your code MUST execute without any errors. 
3. You can add more lines in your code as required.

## Section 1: Data Types and Operations Pt. 1

### Question 1 
**Create the variables with the following composition:**  
1. A vector containing each letter of your first name as its elements.  
2. A variable that contains your name concatenated from the vector created in (1)  
3. A variable containing a sequence from 100 to 120.  
4. Create a matrix of 3x3 dimensions that contains the even sequence of numbers starting from 2.  
5. Assign names to the variables.  


```{r }
#### Start solution ####
```

myFirstName <- c("M", "u", "h", "a", "m", "m", "a", "d")

```{r }

```

mySecondName <- "Shahid"
myFName <- paste(myFirstName, collapse = "")

concateName <- c(myFName, mySecondName) 
concateName

```{r }

```

mySeqNum <- c(100:120)
mySeqNum

```{r }

```

matrixDigit <- seq(2,18, by = 2)
myMatrix <- matrix(matrixDigit,nrow = 3, ncol = 3)
myMatrix

```{r }

```

#### Assign Name To the Variables
name1 <- "Shahid"
name2 = "Mirza"

name1
name2

```{r }
#### End solution ####
```

### Question 2
**Create a factor variable emp_status:**  
1. Containing the categorical variables: Employed, Unemployed, Self-Employed, with each level appearing atleast more than 2.  
2. Display the levels and the factor variable as a table.  
3. Unclass the elements of the factor variable.  


```{r }
#### Start solution ####
```
  
  empCategory<-factor(c("Employed","Unemployed","Self-Employed", "Unemployed","Self-Employed"),levels = c("Employed","Unemployed","Self-Employed"))
  empCategory

```{r }

```

levels(empCategory)
empTab = table(empCategory)
empTab
class(empCategory)

```{r }

```

unclass(empCategory)

```{r }

```


```{r }
#### End solution ####
```

### Question 3
**Create a dataframe object called bank_customers:**  
1. the data frame will have three columns: CustomerID, hasAccount, totalBalance  
2. Fill the data as follows  
    a. Alicia does not have an account. She is here for inquiry   
    b. Nancy is here to check on her account balance of USD 10,000.  
    c. Fernando is here to deposit USD 100 in his account which had no balance.  
    d. Louis will withdraw all his money from the account that had USD 2,000.   
    e. Diane is here for information.   
3. For customers that do not have a value, use NA as placeholder.  
4. Print the number of rows, columns and names for the data frame.  


```{r }
#### Start solution ####
```

dataBankCust <- data.frame(CustomerID = c("Alicia","Nancy","Fernando","Louis","Diane"), hasAccount = c(NA,"Yes","Yes","Yes",NA), totalBalance = c(NA,10000,100,2000,NA) )

dataBankCust

```{r }

```

NumOfRows <- nrow(dataBankCust)
NumOfCol <- ncol(dataBankCust)
dataName<- names(data.frame(dataBankCust))

NumOfRows
NumOfCol
dataName

```{r }

```


```{r }

```


```{r }

```


```{r }
#### End solution ####
```

### Good Job! You have successfully completed the section!

## Section 2: Control Structures 

### Question 1
**Create a variable containing a sequence of numbers from 1 to 100:**  
1. Iterate over the variables and print those numbers which are prime.  

**Create a variable var with a value of 1:**  
1. Create a while loop and increase the value of var at each iteration.   
2. When you reach the 10th prime number, terminate the loop and print the number.  


```{r }
#### Start solution ####
```

numSeq <- c(1:100)
getPrime<-c()

for(x in numSeq){
  if (x >= 2){
    if(any(numSeq == x)){
      fPrime = ( (numSeq && x ) != 0 )
      ##if( (numSeq %% x ) != 0 )
        a = c(numSeq[fPrime], x)
      getPrime <- c(getPrime, x)
      if(length(getPrime) >= 10 ) break
    }
  }
}

getPrime


```{r }

```


```{r }

```


```{r }

```


```{r }

```


```{r }

```


```{r }
#### End solution ####
```

### Question 2
**Create a matrix of size 3x3 called mat_1:**  
1. Iterate over all the values one by one and print the element as well as the position in the matrix (row, col)


```{r }
#### Start solution ####
```

myMatrix<-matrix(1:9,nrow=3,ncol=3)
myMatrix
for(y in 1:ncol(myMatrix))
  for(x in 1:nrow(myMatrix))
    print(paste("Row:",x," Column:",y, " = ", myMatrix[x, y] ))


```{r }

```


```{r }

```


```{r }

```


```{r }

```


```{r }

```


```{r }
#### End solution ####
```

### Good Job! You have successfully completed the section!

## Section 3: Functions

### Question 1
**Create a function called gcd that finds the greatest common divisor of two numbers a and b:**  
1. a and b, should be taken as input.  
2. The function must print the GCD as well as return it.  
3. The output must be saved in a variable called answer.  


```{r }
#### Start solution ####
```

gcd<-function(a, b){
#  a <- 22
#  b <- 11
#  answer <- 0
  if(a < b){
    sNum <- a
  }else{
    sNum <- b
  }
  divSeq <- 1:sNum
  for(x in divSeq){
    if( (a %% x == 0) && (b %% x == 0) ){
      getANS = x
    }
  }
  return(getANS)
}
answer = gcd(22, 11)
print(paste("GCD: ", answer))


```{r }

```

firstNum = as.integer( readline(prompt = "Enter First Number: " ))
secNum = as.integer( readline(prompt = "Enter Second Number: " ))

answer = gcd(firstNum, secNum)
answer

```{r }

```


```{r }

```


```{r }

```


```{r }

```


```{r }
#### End solution ####
```

### Question 2
**Create a function called allConditionsMeet, that checks whether two expressions evaluate to true:**  
1. a and b, should be taken as input.  
2. the function should check if a and b, both conditions, evaluate to True.  
3. The function must returns a boolean value.  
4. Using a method, print the arguments that function takes.  


```{r }
#### Start solution ####
```

allConditionsMeet <- function(a,b){
  
  if( (a < b) && (b > 1) ){
    return (TRUE)
  }else{ return (FALSE) }
  
}

print(allConditionsMeet(8,4))



```{r }

```


```{r }

```


```{r }

```


```{r }

```


```{r }

```


```{r }
#### End solution ####
```

### Good Job! You have successfully completed the section!

## Section 4: Vectorized Operations

### Question 1
**Create two matrices matrix_1 and matrix_2 of dimensions 2x3 and 3x2:**  
1. Perform element-wise multiplication.  
2. Perform matrix multipilcation.   

**Create a 2x2 matrix my_mat:**  
1. Write a function to find the determinant of the matrix.  


```{r }
#### Start solution ####
```

matrix_1 <- matrix(1:6,nrow=2, ncol=3);
matrix_2 <- matrix(1:6,nrow=3, ncol=2);

matrix_1
matrix_2

# Element-Wise Multiplication not Possible (because sise of matrix differ)
# Element-wise possible only if take transpose of any Matrix.

# Multiplication of 2 Matrix are
print( matrix_1 %*% matrix_2 )


```{r }

```

my_mat <- matrix(11:14, nrow=2, ncol=2)
my_mat

getDeterminant <- function(getMatrix){

  myDeterminant <- (( getMatrix[1,1]*getMatrix[2,2] )-( getMatrix[1,2]*getMatrix[2,1] ))
  return(myDeterminant)
}

getAns <- getDeterminant(my_mat)
print(paste("Determinant: ",getAns ))


```{r }

```


```{r }

```


```{r }

```


```{r }

```


```{r }
#### End solution ####
```

### Good Job! You have successfully completed the section!

## Section 5: Date and Time in R

### Question 1
**Use the current date and time and store them into variables curr_date and curr_time respectively: [use sys]**  
1. Convert both into date and time objects using the appropriate functions.  
2. Print the weekday, year, second and hour using the appropriate function and variables.  


```{r }
#### Start solution ####
```

curr_date <- Sys.Date()
curr_time <- Sys.time()
curr_date
curr_time

cDate <- as.POSIXlt(curr_date)
cTime <- as.POSIXlt(curr_time)


getwkDay <- format(cDate,'%d')
getMonth <- format(cDate,'%B')
getYear <- as.numeric(format(cDate,'%Y'))

getHours <- format(cTime,'%H')
getMints <- format(cTime,'%M')
getSeconds <- format(cTime,'%S')


print(paste("Day: ", getwkDay))
print(paste("Month: ", getMonth))
print(paste("Year: ", getYear))

print(paste("Hours: ", getHours))
print(paste("Minutes: ", getMints))
print(paste("Seconds: ", getSeconds))


print(paste("My Date: ", getwkDay, "-" , getMonth, "-" , getYear))
print(paste("My Time: ", getHours, ":" , getMints, ":" , getSeconds))



```{r }

```



```{r }

```


```{r }

```


```{r }

```


```{r }

```


```{r }
#### End solution ####
```

### Question 2
**Create a variable to store current date/time**  
1. Create another variable that stores and set the timezone as GMT-5  
2. Find the difference between your current time and the GMT-5 timezone.  


```{r }
#### Start solution ####
```

myCurr_time <- as.POSIXct(Sys.time())
myCurr_time


GMT_time <- as.POSIXct(myCurr_time, tz="GMT" )
GMT_time

print (GMT_time - curr_time )


```{r }

```


```{r }

```


```{r }

```


```{r }

```


```{r }

```


```{r }
#### End solution ####
```

### Good Job! You have successfully completed the section!

## Section 6: Loop Functions

### Question 1
**Create a function to calculate mean and standard deviation of the provided data**
1. Create a sequence of numbers from 100 to 150 store in a variable called numbers.
1. Use lapply, sapply, apply and tapply to implement the functions on "numbers" [only on the second half of the sequence for tapply]


```{r }
#### Start solution ####
```

numbers <- list(100:150)
numbers

sdL <- lapply(numbers, sd)
sdS <- sapply(numbers, sd)
meanL <- lapply(numbers, mean)
meanS <- sapply(numbers, mean)

print("------ STANDARD DEVIATION ------")
print(paste("LAPPLY (SD): ", sdL))
print(paste("SAPPLY (SD): ", sdS))

print(paste("LAPPLY (MEAN): ", meanL))
print(paste("SAPPLY (MEAN): ", meanS))

mat <- matrix(101:150, 5,10)
print(paste("My Matrix: ",mat))

apply_mean_row <- apply(mat,1,mean)
apply_mean_col <- apply(mat,2,mean)
apply_SD_row <- apply(mat,1,sd)
apply_SD_col <- apply(mat,2,sd)

apply_mean_row
apply_mean_col

apply_SD_row
apply_SD_col


tNumbers <- c(101:150)
tfactor <- gl(5,10)
tNumbers
tfactos

TAPPLY_MEAN = tapply(tNumbers, tfactor, mean)
TAPPLY_SD = tapply(tNumbers, tfactor, sd)
TAPPLY_MEAN
TAPPLY_SD


```{r }

```


```{r }

```


```{r }

```


```{r }

```


```{r }

```


```{r }
#### End solution ####
```

### Question 2
**Create a matrix of dimensions 4x4**
1. Find the row-wise and column-wise mean of the matrix.
2. Print the values.


```{r }
#### Start solution ####
```

S6Q2Mat <- matrix(1:16,4,4)
S6Q2Mat

apply(S6Q2Mat, 1, mean) #ROW-WISE MEAN OF MATRIX
apply(S6Q2Mat, 2, mean) #COLUMN-WISE MEAN OF MATRIX


```{r }

```


```{r }

```


```{r }

```


```{r }

```


```{r }

```


```{r }
#### End solution ####
```

### Good Job! You have successfully completed the section!

## Section 7: Data Split

### Question 1
**Using the data frame Orange:**  
1. Using split function to break down the dataset on circumference and store it in 'split_data' variable.   
2. Print the values for split_data where circumference is 30 and 75.  
3. Find the average age of the tree when the circumference is 30 and when circumference is 214.  


The dataset is loaded and the variable Orange contains the respective dataset.  


```{r }
library(datasets)
```


```{r }
head(Orange)
```


```{r }
#### Start solution ####
```

split_data <- split(Orange, Orange$circumference)

split_data$'30'
split_data$"75"

age1_data <- split_data$'30'
age2_data <- split_data$'214'

age1_data
age2_data

mean(age1_data$age)
mean(age2_data$age)


```{r }

```


```{r }

```


```{r }

```


```{r }

```


```{r }

```


```{r }
#### End solution ####
```

### Save it and push .Rmd and .html file to your Github Repository
## GOOD JOB! You have successfully finished the course!